---
title: "Coffee Data Exploration"
author: "Ted Laderas"
description: Understanding coffee production and consumption across the world.
categories: tidytuesday
date: "7/8/2020"
output: 
  distill::distill_article:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


# What was your dataset?

Load your dataset in with the function below. The input is the date the dataset was issued. You should be able to get this from the `tt_available()` function.

```{r}
coffee <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')

```

# What was your question?

Given your inital exploration of the data, what was the question you wanted to answer?

Does processing method affect overall coffee rating?

# Initial Skim of Data

```{r}
skimr::skim(coffee)
```


# Total Number of Samples Per Country

```{r}
coffee %>%
  janitor::tabyl(country_of_origin) %>%
  arrange(desc(n)) %>%
  gt::gt()

```


# Distribution of `total_cup_points` versus `processing_method`

```{r}
ggplot(coffee) + 
  aes(y=total_cup_points, x=processing_method, fill=processing_method) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=90, hjust = 1)) +
  coord_flip() 
```

# Counts of Country of Origin versus Processing Method

```{r fig.height=8}
coffee %>%
  mutate(country_of_origin= fct_rev(country_of_origin)) %>%
  ggplot() +
  aes(y=country_of_origin, x=processing_method, 
      color=processing_method) +
  geom_count() +
  theme(axis.text.x = element_text(angle=90))
```

Here's a sortable table of the above table

```{r}
library(reactable)
coffee %>%
  janitor::tabyl(country_of_origin, processing_method) %>%
  reactable::reactable()
```


# Sorted Heatmap of scores by `total_cup_points`

```{r}
coffee %>% mutate(sample_id = rownames(coffee)) %>%
  select(sample_id, country_of_origin, total_cup_points, aroma, flavor, acidity, body, balance, uniformity, clean_cup, sweetness, cupper_points)%>%
  pivot_longer(cols = c(aroma, flavor, acidity, body, balance, uniformity, clean_cup, sweetness, cupper_points), names_to="type", values_to="score") %>%
  mutate(sample_id = fct_reorder(sample_id, total_cup_points)) %>%
  ggplot() +
  aes(y=sample_id, x=type, fill=score) +
    geom_tile()
```

# Bi-clustered Heatmap of Scores

```{r}
library(heatmaply)

coffee %>% mutate(sample_id = rownames(coffee)) %>%
  select(aroma, flavor, acidity, body, balance, uniformity, clean_cup, sweetness, cupper_points) %>% heatmaply()
```

## Processing Method: `Dry / Natural`

```{r}
coffee %>%
  filter(processing_method == "Natural / Dry") %>%
  mutate(country_of_origin = fct_reorder(country_of_origin, total_cup_points, median)) %>%
ggplot() + 
  aes(y=total_cup_points, x=country_of_origin, fill=country_of_origin) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=90), legend.position = "none") +
  coord_flip() +
  labs(title="Tanzania leads with ratings in Natural/Dry")
```

## Processing Method: `Washed / Wet`

```{r}
coffee %>%
  filter(processing_method == "Washed / Wet") %>%
  mutate(country_of_origin = fct_reorder(country_of_origin, total_cup_points, median)) %>%
ggplot() + 
  aes(y=total_cup_points, x=country_of_origin, fill=country_of_origin) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle=90), legend.position = "none") +
  coord_flip() +
  labs(title="US leads in Ratings in Washed/Wet")
```

## Mexico: Processing Methods

```{r}
coffee %>%
  filter(country_of_origin == "Mexico") %>%
  mutate(processing_method = fct_reorder(processing_method, total_cup_points, median)) %>%
  ggplot() +
  aes(y=total_cup_points, x=processing_method,  fill=processing_method) +
  geom_boxplot(color="black") +
  coord_flip()
  

```

# Linear model of `total_cup_points`

```{r}
broom::tidy(lm(total_cup_points ~
                 country_of_origin +
                 category_one_defects, data=coffee)) %>%
  filter(p.value < 0.05) %>%
  arrange(p.value)

```

