---
title: 'group_by/summarize: Not just for numeric values!'
author: ~
date: '2018-05-24'
slug: group-by-summarize-not-just-for-numeric-values
categories: 
  - tidyverse
tags: []
header:
  caption: ''
  image: ''
description: |
   Learn some more about the many uses of `group_by()`/`summarize()`.  
output:
  distill::distill_article:
    self_contained: true
---
```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
library(DT)
```


Even though I've been using the `tidyverse` for a couple of years, there's always a couple new applications of `tidyverse` verbs.

This one, in retrospect, is pretty simple. I had a one to many table that I wanted to collapse, tidy-style. Let's look at the `diamonds` dataset:

```{r}
diamonds %>% select(color, cut) %>%
  head() %>%
  knitr::kable()
```

What if we wanted to collapse all the entries for each color into a single line? There's 7 different colors, so we can use a combination of `group_by` on `color` and use the `paste()` function within `summarize()` to get what we want, which I've called `all_colors` here. By specifying the `collapse` argument, we can specify the delimiter within that column: 

```{r}
diamonds %>% select(color, cut) %>% 
  group_by(color) %>% 
  summarize(all_colors=
              paste(cut, collapse=";"))
```

Thanks to Ken Butler, who pointed out that the `tidyverse` way (via `stringr`) is to use `str_c` instead: 

```{r}
diamonds %>% select(color, cut) %>% 
  group_by(color) %>% 
  summarize(all_colors=
              stringr::str_c(cut, collapse=";")) 
```

Finally, if we wanted to just get the unique values of the cuts in a single line, we can use `unique`:

```{r}
diamonds %>% select(color, cut) %>% 
  group_by(color) %>% 
  summarize(all_colors=
              paste(unique(cut), collapse=";")) 
```