---
title: "Tidy Tuesday: Crop Production"
author: "Ted Laderas"
description: Understanding crop production across the world.
categories: tidytuesday
date: "9/3/2020"
output: 
  distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Look at the available datasets

```{r eval=FALSE}
library(tidytuesdayR)
#This will open up in the help window
tidytuesdayR::tt_available()
```

# What was your dataset?

Load your dataset in with the function below. The input is the date the dataset was issued. You should be able to get this from the `tt_available()` function.

```{r}
#incoming data comes in as a list
datasets <- tidytuesdayR::tt_load("2020-09-01")
#show the names of the individual datasets
names(datasets)
```

## Key Crop Yields

```{r}
key_crop_yields <- datasets$key_crop_yields
```


## Visdat

```{r}
visdat::vis_dat(key_crop_yields)

```

## Skimr

```{r}
skimr::skim(key_crop_yields)

```


# What was your question?

Given your inital exploration of the data, what was the question you wanted to answer?

> How have key crop yields changed over time?

# What were your findings?

Put your findings and your visualization code here.

```{r fig.height=14, fig.width=14}
key_crop_yields %>%
  tidyr::pivot_longer(cols = contains("(tonnes"), names_to="crop",
                      values_to="Yield") %>%
  ggplot() + aes(x=Year, y= Yield, group=crop, color=crop) + geom_line() + facet_wrap(~Entity)
```

Let's try and estimate whether a country is increasing its yield or decreaing its yield over time. I'll use `lm()` to run a linear regression on each entity in the data, and use `broom::tidy()` to pull out the estimates of the slopes.

```{r}
model_results <- key_crop_yields %>%
  tidyr::pivot_longer(cols = contains("(tonnes"), names_to="crop",
                      values_to="Yield") %>%
  mutate(crop=str_replace(crop,"\\(tonnes per hectare\\)", "")) %>%
  tidyr::drop_na(Yield) %>%
  nest_by(Entity, crop) %>%
  mutate(num_points = nrow(data)) %>%
  mutate(model=list(lm(Yield ~ Year, data=data))) %>%
  summarize(num_points, broom::tidy(model)) %>%
  filter(term == "Year") %>%
  arrange(Entity, desc(estimate))

model_results
```

Looking at the United States, there are mostly increases in crop yield.

```{r}
model_results %>%
  filter(Entity == "United States") 

```

We can rank the top producers by crop:

```{r}
ranked_by_slope <- model_results %>%
  ungroup() %>%
  group_by(crop) %>%
  summarize(Entity, crop, num_points, estimate=signif(estimate, digits = 3), rank = row_number(desc(estimate))) %>%
  arrange(crop, rank)

ranked_by_slope
```

Finally, let's do histograms by crop:

```{r}
ranked_by_slope %>%
  ggplot() +
  aes(x=estimate, fill=crop) +
  geom_histogram() +
  geom_vline(xintercept = 0, lty =2) +
  labs(title = "Crop productivity across countries (tonnes/hectare/year)",
       subtitle = "Positive values = increase, Negative values = decrease") +
  theme_minimal() +
  theme(legend.position = "none") +
  facet_wrap(~crop, scales = "free")
```

# What did you learn?

Were there any lessons you learned? Any cool packages you want to talk about?
