---
title: "Tidy Tuesday Template"
author: "You!"
date: "3/26/2020"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Look at the available datasets

```{r}
library(tidytuesdayR)
#This will open up in the help window
tidytuesdayR::tt_available()
```

# What was your dataset?

Load your dataset in with the function below. The input is the date the dataset was issued. You should be able to get this from the `tt_available()` function.

```{r}
#incoming data comes in as a list
datasets <- tidytuesdayR::tt_load("2020-09-01")
#show the names of the individual datasets
names(datasets)
```

## Key Crop Yields

```{r}
key_crop_yields <- datasets$key_crop_yields
```


## Visdat

```{r}
visdat::vis_dat(key_crop_yields)

```

## Skimr

```{r}
skimr::skim(key_crop_yields)

```


# What was your question?

Given your inital exploration of the data, what was the question you wanted to answer?

> How have key crop yields changed over time?

# What were your findings?

Put your findings and your visualization code here.

```{r}
key_crop_yields %>%
  filter(Entity %in% c("United States", ))
  tidyr::pivot_longer(cols = contains("(tonnes"), names_to="crop",
                      values_to="Yield") %>%
  ggplot() + aes(x=Year, y= Yield, group=crop, color=crop) + geom_line() + facet_wrap(~Entity)
```

```{r}
key_crop_yields %>%
  tidyr::pivot_longer(cols = contains("(tonnes"), names_to="crop",
                      values_to="Yield") %>%
  tidyr::drop_na(Yield) %>%
  nest_by(Entity, crop) %>%
  mutate(model=list(lm(Yield ~ Year, data=data))) %>%
  summarize(broom::tidy(model)) %>%
  filter(term == "Year") %>%
  arrange(Entity, desc(estimate))

```


# What did you learn?

Were there any lessons you learned? Any cool packages you want to talk about?
