---
title: "Registered Nurses in the United States and Territories"
author: "Ted Laderas"
description: Understanding wages for Registered Nurses.
date: "10/5/2021"
categories: tidytuesday
tags: [tidytuesday]
output: 
  distill::distill_article:
    self-contained: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(bplots)
library(heatmaply)
```

# Research Question(s)

1. Which states have the highest overall wages for registered nurses? When did this happen?
1. Have wages increased overall for registered nurses across all states? 

# Loading Data

We'll use the Tidy Tuesday code to directly load the data from the GitHub repository. We'll also pass it into `janitor::clean_names()` to standardize the column names. (Life is too short to have to worry about whitespace and capitalization.)

```{r}
nurses <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-05/nurses.csv') %>% janitor::clean_names()
```

# Initial EDA

We can see there are 22 columns overall. 21 of these are numeric. 

```{r}
skimr::skim(nurses)
```

```{r}
head(nurses)
```
Looking at how years are divided.

```{r}
nurses %>%
  count(year)
```

Hmmm. 54 entries per year. This includes: D.C., Virgin Islands, Puerto Rico, and Guam in addition to the 50 states.

```{r}
nurses %>%
  count(state)
```

The mean total number of nurses overall states shows an upward trend, except for a blip in 2012-2013.

```{r}
nurses %>%
  group_by(year) %>%
  summarize(mean_employed_rn = mean(total_employed_rn, na.rm=TRUE)) %>%
  ggplot() +
  aes(x=year, y=mean_employed_rn) %>%
  geom_line()
```
Let's visualize whether hourly wages are increasing or decreasing across the dataset by making a heatmap. On the x-axis, we will visualize `year`, and we will visualize by `state` on our y-axis. We're going to map the `fill` value to `hourly_wage_median`:

```{r}
nurses %>%
  mutate(state=forcats::fct_rev(state)) %>%
  ggplot() +
  aes(x=year, y=state, fill=hourly_wage_median) +
  geom_tile()
```

## Scaling the data by `state`

Looking for trends in the `nurses` data, let's try and scale each income so we can emphasize whether there were increases or decreases within each state. We're just looking for trends here and whether the slope of these trends is the same for each state.

Note that by scaling within a state (transforming each value to a z-score), we are losing information, but we can see whether wages are steadily increasing for each of the states/territories.

In general, with some exceptions (Guam and Virgin Islands), most registered nurses saw an increase in median hourly wages from 1998 to 2020.

```{r fig.width=6, fig.height=8}
nurses %>%
  mutate(state=forcats::fct_rev(state)) %>%
  group_by(state) %>%
  mutate(scaled_income = scale(hourly_wage_median)) %>%
  ggplot() +
  aes(x=year, y=state, fill=scaled_income) +
  geom_tile(color="grey10") +
  scale_fill_distiller() +
  bplots::theme_avenir()
```

Since we looked at median hourly income, the question is whether these trends are the same or different for the 10th and 90th percentiles of registered nurses.

## 10th Percentile

```{r fig.width=6, fig.height=8}
nurses %>%
  mutate(state=forcats::fct_rev(state)) %>%
  group_by(state) %>%
  mutate(scaled_income = scale(hourly_10th_percentile)) %>%
  ggplot() +
  aes(x=year, y=state, fill=scaled_income) +
  geom_tile(color="grey10") +
  scale_fill_distiller() +
  bplots::theme_avenir() +
  theme(axis.text.x=element_text(angle=90))
```

## 90th Percentile

For the most part, if you are in the 90th percentile of hourly wages, you have seen a leveling off of income after about 2008. After 2008, the 90th income seems pretty static and unchanging.


```{r fig.width=6, fig.height=8}
nurses %>%
  mutate(state=forcats::fct_rev(state)) %>%
  group_by(state) %>%
  mutate(scaled_income = scale(hourly_90th_percentile)) %>%
  ggplot() +
  aes(x=year, y=state, fill=scaled_income) +
  geom_tile(color="grey10") +
  scale_fill_distiller() +
  bplots::theme_avenir() +
  ggtitle("90 percentile RNs have slower increases in income than the 10%")
```

# Making heatmaps with dendrograms

## Pivoting the data to be wider

One question we might ask are whether there are groupings by states in terms of the wage increases. 

We can do this by pivoting the data and using the `{heatmaply}` package to make a matrix input suitable for `heatmaply::heatmaply()`.

Here, we take `hourly_wage_median` and use it in the values of our matrix. Our rows correspond to `state` and our columns correspond to `year`.

```{r}
nurse_median_frame <- nurses %>%
  select(state, year, hourly_wage_median) %>%
  arrange(year) %>%
  tidyr::pivot_wider(names_from = year, values_from = hourly_wage_median) 

nurse_median_matrix <- nurse_median_frame[,-1]
rownames(nurse_median_matrix) <- nurse_median_frame[["state"]]
nurse_median_matrix <- as.matrix(nurse_median_matrix)

head(nurse_median_matrix)
```

## Heatmap with No scaling

We can now ask questions about the actual income values. We make `heatmaply` only look at computing a dendrogram for the rows (states) to look for clustering patterns. 

Note we have to set our `scale` argument to `none` here. 

```{r fig.width=6, fig.height=8}
heatmaply(nurse_median_matrix, dendrogram = "row", 
          Colv = c(1:23), scale="none",
          main = "Oregon, California, and Hawaii have the highest median wage from 2017-2020")
```

## Scaling by `state`

If we are interested in relative (scaled) values, the dendrogram is a little less interesting. Overall you can see that all states showed an increase in hourly median wage over the years. 

```{r fig.width=6, fig.height=8}
heatmaply(nurse_median_matrix, dendrogram = "row", 
          Colv = c(1:23), scale="row", 
          main="Upward trends overall in terms of hourly median wage")
```

# Conclusions

This was a nice dataset to get back into Tidy Tuesday. 

- Median wages have increased across all states for Registered Nurses.
- Hawaii, Oregon, and California have the highest overall wages for Registered Nurses