---
title: "Tidy Tuesday R packages"
---

```{r}
tuesdata <- tidytuesdayR::tt_load('2023-12-26')
## OR
tuesdata <- tidytuesdayR::tt_load(2023, week = 52)

cran_20221122 <- tuesdata$cran_20221122
external_calls <- tuesdata$external_calls
internal_calls <- tuesdata$internal_calls
```

```{r}
library(tidyverse)
library(tidygraph)
library(forcats)
```

## What are the top packages by centrality?

I'm actually really surprised that the top ranking package by centrality is `{duckdb}`. It's really become very ubiquitous.

```{r}
cran_20221122 |> dplyr::filter(centrality_dir_mn_no0 > 2000) |> 
  mutate(package=fct_reorder(package, centrality_dir_md_no0)) |> 
  slice_max(centrality_dir_md_no0, n=20) |> 
  ggplot() + aes(x=package, y=centrality_dir_md_no0) +
  geom_col() + theme(axis.text = element_text(angle = 0)) + coord_flip()
```

```{r}
cran_20221122 |> #dplyr::filter(centrality_dir_mn_no0 > 2000) |> 
  mutate(package=fct_reorder(package, node_degree_max)) |> 
  slice_max(node_degree_max, n=20) |> 
  ggplot() + aes(x=package, y=node_degree_max) +
  geom_col() + theme(axis.text = element_text(angle = 0)) + coord_flip()
```

## Centrality Measures

```{r}
my_graph <- cran_20221122 |>
  ggplot() +
  aes(x= centrality_dir_md_no0, y=centrality_dir_md, package=package) +
  geom_point()

plotly::ggplotly(my_graph)
```

The undirected versus directed measures are less correlated, indicating there are some packages that have high undirected centrality but not high directed centrality.

```{r}
my_graph <- cran_20221122 |>
  ggplot() +
  aes(x= centrality_undir_md_no0, y=centrality_dir_md_no0, package=package) +
  geom_point()

plotly::ggplotly(my_graph)
```

## Documentation versus Centrality

```{r}
my_graph <- cran_20221122 |>
  dplyr::filter(centrality_dir_md_no0 < 1000) |>
  dplyr::filter(docchars_per_par_exp_md < 1000) |>
  ggplot() +
  aes(x=docchars_per_par_exp_md, y=centrality_dir_mn_no0, package=package) +
  geom_point()

plotly::ggplotly(my_graph)
```

