---
title: 'Session 1: Learning ggplot2'
author: "Ted Laderas"
date: "7/13/2017"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## These Slides are Here:

http://laderast.gihub.io/RBoot1/

## If you're new to R

We're going to start from the very beginning. 

We want to show you what's possible first, and then how to get there.

Remember: this is not easy material, but be patient and you can do it!

## If you've used R before

The `tidyverse` simplifies a lot of the data processing steps needed to get data into statisical procedures. 

It's really different than base-R, but it makes a lot of everyday procedures simple.

## Planned Sessions

- July 14 - `ggplot2` - getting to visualization
- July 28 - `dplyr` - cleaning and shaping your data
- August 11 - `tidyr` - the philosophy of tidy data
- August 25 (tenative) - `broom` - doing simple statistics in the tidyverse way 

## DataCamp

DataCamp encourages experimentation! Don't be afraid of it!

- Script Window - CTRL+ENTER (PC) / CMD+ENTER (Mac) - submit final answer here
- Console Window - can experiment here!

## What is a `data.frame`?

- Table with Columns
- Each column is called a variable
    + Each variable can be a different type:
    + Numeric, Categorical (`factor`), String (`character`)
- Each row is an observation

```{r echo=TRUE}
data(iris)
head(iris)
```

## Looking at a single variable

```{r echo=TRUE}
iris$Sepal.Length
length(iris$Sepal.Length)
```

## Let's Get Started!

+ You'll need to get registered at DataCamp first: http://datacamp.com
+ Datacamp course: http://datacamp.com/courses/rbootcamp 
+ Be patient and curious! Try things out before you answer
+ Help each other! You can do it!
+ Use Post-its to signal for help (red - need help, green - I'm ok)
+ Online students: post questions in chat window, we'll do our best to answer

## Other forms of This Course

- Original Workshop form: https://github.com/Cascadia-R/gRadual-intRoduction-tidyverse
- Bookdown form (good for reference): https://tidyverse-intro.github.io